---
title: "hw6"
output: pdf_document
date: "2023-10-23"
---
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d)) 

library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```
My function needs to input PDB data and output a plot for the specified
protein

```{r}
Jackie_function <- function(x) {
  protein <-read.pdb(x) # read file
  chain <- trim.pdb(protein, chain = "A", elety = "CA")
  b_factors <- chain$atom$b #extracting b factors 
  plotb3(b_factors,sse=bfactors.chain, typ="l",ylab="Bfactor") #plot- 
}

```
This function automates the above example process so you can just input the structure file (PBD) and get out the same plotting information, but it will be quicker than typing everything out like in the example. 
For any protein input into function- output is a plot with:
x axis= residue #
y axis= b factors