---
title: "halloween-mini-project"
format: pdf
---

```{r}

candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)


```

Q1. How many different candy types are in this dataset?
```{r}
nrow(candy)

```

Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)

```
Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
```{r}
candy["Reese's Peanut Butter cup", ]$winpercent


```

Q4. What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat", ]$winpercent


```

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent


```
```{r}
#install.packages("skimr")
#library("skimr")
#skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
Yes, winpercent is on a different scale

Q7. What do you think a zero and one represent for the candy$chocolate column?
If it is a chocolate-y candy or not- 0= not, 1= yes

Q8. Plot a histogram of winpercent values
```{r}
library(ggplot2)
ggplot(candy,aes(x=winpercent))+
  geom_histogram()
```
Q9. Is the distribution of winpercent values symmetrical?

No

Q10. Is the center of the distribution above or below 50%?
 Below

Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$nougat)])
mean(candy$winpercent[as.logical(candy$chocolate)])
mean(candy$winpercent[as.logical(candy$fruity)])

```
Higher!

Q12. Is this difference statistically significant?
```{r}
choc<-candy$winpercent[as.logical(candy$chocolate)]
fruit<-candy$winpercent[as.logical(candy$fruity)]
t.test(choc,fruit)
```
Yes!

Q13. What are the five least liked candy types in this set?
```{r}
head(candy[order(candy$winpercent, decreasing = TRUE), ], n = 5)
head(candy[order(candy$winpercent),], n=5)

```
Nik L Nip, Boston Baked Beans, Super Bubble, Jawbusters
Q14. What are the top 5 all time favorite candy types out of this set?
Reese's Peanut Butter cup, Reese's miniatures, Twix, Kit Kat, Snickers

Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
ggplot(candy)+
  aes(x=winpercent, y=rownames(candy)) +
  geom_col()
```
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 


```
Now, for the first time, using this plot we can answer questions like:
- Q17. What is the worst ranked chocolate candy?
Boston Baked Beans

- Q18. What is the best ranked fruity candy?
Starburst
```{r}
#install.packages("ggrepel")
#library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) #+ 
  #geom_text_repel(col=my_cols, size=3.3, max.overlaps = 10)
```
Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
Tootsie Roll Midgies

Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
Nik L Nip (least popular), Nestle Smarties, Ring Pop, Mr. Good Bar, Hershey's Milk Chocolate 

Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().
```{r}
ggplot(candy, aes(pricepercent,reorder(rownames(candy),winpercent)))+
  geom_col()
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
#install.packages("corrplot")
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```
Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
chocolate and fruity
Q23. Similarly, what two variables are most positively correlated?
```{r}
cor(candy$chocolate,candy$winpercent)
cor(candy$chocolate,candy$bar)
```
winpercent and chocolate

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}

#plot(pca$x[,1:2])
plot(pca$x[,1:2], col=my_cols, pch=16)

```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p

```

```{r}
p + #geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

```


```{r}
#install.packages("plotly")
#library(plotly)
#ggplotly(p)
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

Q24
Pluribus and fruity
