---
title: "Class 5: Data Visualization"
author: "Jackie"
format: pdf
---

##Using GGplot
#if you don't have a library, download it by using 'install.packages(')


```{r}
library(ggplot2)
ggplot(cars)


ggplot(cars) +
  aes(x=speed, y=dist)
```  
As we can see above we don’t have any points in our plot yet. Adding them is the job of the geom_point() function discussed next.
```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() + 
  geom_smooth(method="lm", se=FALSE) +
  theme_bw() +
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed (MPH)", 
       y="Stopping Distance (ft)",
       subtitle = "Speed vs distance",
       caption="Dataset: 'cars'")
  
```  
#GENES DATAset  
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes) 

nrow(genes)
ncol(genes)
table(genes$State)


p<- ggplot(genes) +
aes(x=Condition1, y=Condition2,col=State) +
geom_point()

p+ scale_colour_manual( values=c("blue","black","red") )
```
## 7 going further

 File location online
```{r}

url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)

library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

gapminder_2007_gg<-ggplot(gapminder_2007)+
  aes(x=gdpPercap,y=lifeExp,color=continent,size=pop)+
  geom_point(alpha=.5)+
   scale_size_area(max_size = 10)
```
#2007 vs 1957
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)

gapminder_1957_gg<- ggplot(gapminder_1957)+
  aes(x=gdpPercap,y=lifeExp,color=continent,size=pop)+
  geom_point(alpha=.7)+
   scale_size_area(max_size = 10)


gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 8) +
  facet_wrap(~year)
```
#BAR CHARTS

```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop, fill=gdpPercap))

ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=gdpPercap) +
  geom_col()
```
```{r}

ggplot(gapminder_top5) +
  aes(x=reorder(country, -pop), y=pop, fill=country) +
  geom_col(col="gray30") +
  guides(fill="none")
```

```{r}

head(USArrests)
USArrests$State <- rownames(USArrests)
ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_col() +
  coord_flip()

ggplot(USArrests) +
  aes(x=reorder(State,Murder), y=Murder) +
  geom_point() +
  geom_segment(aes(x=State, 
                   xend=State, 
                   y=0, 
                   yend=Murder), color="blue") +
  coord_flip()
