---
title: "Class07"
format: pdf
---

```{r}
rnorm(10)
hist(rnorm(10000,mean=3))
```

```{r}

tmp<-c(rnorm(30,3),rnorm(30,-3))
x<- cbind(x=tmp,y=rev(tmp))
plot(x)
```


```{r}

#k<- kmeans(x, centers,iter.max=10L...)
k<-kmeans(x,centers=2,nstart=20)
k
```

> Q. How many points are in each cluster

```{r}
k$size

``` 
<Q2. The clustering result i.e. membership vector?

```{r}
k$cluster

```


>Q4. Make a plot of our data colored by clustering results with optionally the cluster centers shown

```{r}

plot(x,col=k$cluster)
points(k$centers,col="blue",pch=15,cex=2)
```
>Q5. Run Kmeans again but cluster into 3 groups and plot the results

```{r}
k<-kmeans(x,centers=3,nstart=20)
k$cluster
plot(x,col=k$cluster)

```

#Hierarchical Clustering

The main function in baseR is 'hclust()'

```{r}

hc<-hclust(dist(x))

plot(hc)
abline(h=8)
```

```{r}
grps<-cutree(hc,h=8)
plot(grps)

```

>Q. Plot our hclust rsults in terms of our data colored by cluster membership

```{r}

plot(x,col=grps)

```
Lab portion
```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names=1)
head(x)
dim(x)
```
Q1> 5 columns and 17 rows, but if you do row.names=1, there are 17 rows and 4 columns

Q2> I prefer to import the df using row.names=1 so I know I am always working with the same df in following lines
`
```{r}
colors= (c("orange","red","blue","dark green"))
barplot(as.matrix(x), beside=F, col=colors)


```
Q5> horiz! I"f FALSE, the bars are drawn vertically with the first bar to the left. If TRUE, the bars are drawn horizontally with the first at the bottom."


```{r}
pairs(x, col=colors, pch=16)



```
Q6> There is a stronger outlier in the N. Ireland dataset that may affect the coorelation value


```{r}
t(x) #transpose x
pca<- prcomp(t(x))
summary(pca)
```
```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))


```


Q8> Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.
```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),col=colors)


```

```{r}
#variance from diff PC
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
## or the second row here...
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )

```

Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 maninly tell us about?
```{r}

##PC2
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```
It tells us about Fresh_potatoes and soft_drinks. wales eats more fresh potatoes than the rest and drinks fewer soft drinks.

Q10> How many genes and samples are in this data set?
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
dim(rna.data)
```
